---
title: "Basic R Programming"
format: html
editor: visual
---

ST 558

Ella Gruchacz

# Task 1: Basic Vector Practice

Create two vectors, one for blood pressure pre treatment and one post treatement.
```{r}
pre_bp <- c(120, 151, 125, 126, 115, 132, 132, 129, 134, 139, 127, 122, 127, 135, 133, 128, 147, 138, 140, 132)

post_bp <- c(127, 145, 135, 122, 115, 122, 123, 126, 126, 129, 132, 146, 120, 114, 121, 120, 128, 120, 133, 115)
```

Assign names to each vector element.
```{r}
names(pre_bp) <- paste("Subject", 1:20, sep = "_")
names(post_bp) <- paste("Subject", 1:20, sep = "_")
```

Create a vector storing the difference in BP measurement between post treatment and pre treatment, and calculate the average decrease in blood pressure across all patients.
```{r}
bp_diff <- post_bp - pre_bp
mean(bp_diff)
```
Find which subjects had a positive change, their BP was lower after treatment.
```{r}
which(bp_diff < 0)

```

Save the difference in BP measurements for subjects that experienced a positive change.
```{r}
bp_dec <- bp_diff[bp_diff < 0]
```

Find the average difference in BP for subjects that experienced a positive change.
```{r}
mean(bp_dec)
```
# Task 2: Basic Data Frame Practice

Create a data frame corresponding to the data patient, pre_bp, post_bp, and diff_bp.

```{r}
patients <- names(pre_bp)
bp_data <- data.frame(patients, pre_bp, post_bp, bp_diff, row.names = NULL)
```

Return the rows of the data frame where the diff_bp column is negative.
```{r}
bp_data[bp_data$bp_diff < 0,]
```
Create a new column of the data frame corresponding to TRUE if the post_bp is less than 120 and FALSE if it is not.
```{r}
bp_data$less_120 <- bp_data$post_bp < 120
```

Print the data frame.
```{r}
knitr::kable(bp_data)

```

# Task 3: List Practice

Recreate the last data frame with new data.

```{r}
placebo <- data.frame(
                      patients = paste("Subject", 1:20, sep = "_"), 
                      pre_trt = c(138, 135, 147, 117, 152, 134, 114, 121, 131, 130),
                      post_trt = c(105, 136, 123, 130, 134, 143, 135, 139, 120,124))
placebo$trt_diff <- placebo$post_trt - placebo$pre_trt
                      
placebo$less_120 <- placebo$trt_diff < 120
```

Create a list of two data frames.
```{r}
exper <- list(treatment = bp_data, placebo = placebo)
```

Access the first element in three ways.

* $ operator
```{r}
exper$treatment
```
* Single square brackets []
```{r}
exper[1]
```
* Double square brackets [[]]
```{r}
exper[[1]]
```

 In one line of code, access the pre_bp column of the placebo data frame.
```{r}
exper$placebo$pre_trt
```

